---
title: "NetCOLOR project"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    highlight: zenburn
    css: "theme.css"
    toc: true
    toc_float:
      collapsed: false
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "center",
  cache = TRUE
)

library(here)
library(tidyverse)
library(kableExtra)

library(ggpmthemes)

# Set default ggplot2 font size and font family
# devtools::install_github("PMassicotte/ggpmthemes")
theme_set(theme_light_modified(
  base_family = "Montserrat Alternates",
  base_size = 10
))

theme_update(
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  strip.background = element_blank(),
  strip.text = element_text(face = "bold", size = 10, color = "#4c4a4a")
)

source(here("R/zzz.R"))
```

```{r open_data}
hplc <- vroom::vroom(here("data/clean/hplc.csv"))
metadata <- vroom::vroom(here("data/clean/metadata.csv"))
absorption <- vroom::vroom(here("data/clean/absorption.csv"))
```

This section shows some graphics about the processed data. The data consists in three `.csv` files:

1.  `metadata.csv`

2.  `hplc.csv`

3.  `absorption.csv`

`sample_id` *438116* was duplicated and removed from the data.

## Notes

- Will be using `depth` instead of `pressure` to characterize the depth of the measurements.

## Mismatch between metadata, HPLC and absorption

Now, this table shows the metadata that have no HPLC values associated.

```{r}
metadata %>% 
  anti_join(hplc, by = "sample_id") %>% 
  rmarkdown::paged_table()
```

These are the absorption spectra with no HPLC association.

```{r}
absorption %>% 
  anti_join(hplc, by = "sample_id") %>% 
  distinct(sample_id, .keep_all = TRUE) %>% 
  rmarkdown::paged_table()
```

## First visualizations

Most of the data is from the Scotian Shelf, Labrador Sea and Scotian Shelf and Slope.

```{r}
knitr::include_graphics(pdf2png(here("graphs/06_number_measurements_cruise.pdf")))
```

The number of observation is also relatively stable through the years.

```{r}
knitr::include_graphics(pdf2png(here("graphs/06_number_measurements_per_year.pdf")))
```

## Bioregions

Observations were classified into bioregions based on the following criteria (where `position` is `south` if `latitude < 48`):

```{r, eval=FALSE, echo=TRUE}
bathymetry > -300 & position == "North" ~ "Labrador & Greenland Shelves (LGS)"
bathymetry <= -300 & position == "North" ~ "Labrador Sea Bassin (LSB)"
bathymetry >= -300 & position == "South" & yday <= 180 ~ "Scotian Shelf Spring (SSSp)"
bathymetry >= -300 & position == "South" & yday > 180 ~ "Scotian Shelf Fall (SSFa)"
bathymetry < -300 & position == "South"~ "Northwest Atlantic Bassin ocean (NAB)"
```

```{r, fig.align='center'}
knitr::include_graphics(pdf2png(here("graphs/04_number_observations_per_bioregion.pdf")))
```

## Geographical map

This is not the final map, but it gives a first spatial overview of the data.

```{r, fig.align='center', fig.cap="kjljk"}
knitr::include_graphics(pdf2png(here("graphs/fig01.pdf")))
```

## HPLC

-   I also noticed there are a lot of `0` in the HPLC data. After discussions, they are true `0` and not missing values.

```{r}
hplc %>%
  select(-abs_vol_l) %>%
  pivot_longer(chlb:zea) %>%
  drop_na() %>%
  mutate(is_zero = value == 0) %>% 
  count(name, is_zero) %>%
  ggplot(aes(x = n, y = name, fill = is_zero)) +
  geom_col() +
  paletteer::scale_fill_paletteer_d("ggthemes::wsj_colors6") +
  labs(
    y = NULL,
    fill = "Value\nis zero"
  ) 
```

- Some HPLC `sample_id` were not numeric (ex.: FL002), I removed them from the data.

-   There are `hplchla` and `hplcchla` variables in the data. I have merge both columns into one because I suspect it is an error and should be the same variable.

- `but19` was summed as `but19` + `butlike`.

- `hex19` was summed as `hex19` + `hexlike` + `hexlike2`.

- I have calculated `aphy_specific` using `hplcchla`.

## Absorption

-   Here is how I named absorption data:

    -   `anap` is the data contained in files with `detritus` in their name.
    -   `ap` is the data contained in files with `particulate` in their name.
    -   `aphy` is the data contained in files with `phytoplankton` in their name.


- Absorption spectra with any values <= 0 between 350 and 400 nm have been removed.

- Absorption spectra with `aphy(440)` < `aphy(410)` have been removed (possible problem with pigment extraction).

Overview of `aphy` per bioregion. We can see that there are some negative values at low wavelengths (see `Northwest Atlantic Bassin ocean (NAB)`).

```{r}
knitr::include_graphics(pdf2png(here("graphs/07_aph_spectral_profiles_per_bioregion.pdf")))
```

### Calculating snap

> The fit was done for data between 380 and 730 nm, excluding the 400--480 and 620--710 nm ranges to avoid any residual pigment absorption that might still have been present after sodium hypochlorite treatment [@Babin2003].

A R2 of 0.90 was used to filter out bad models.

```{r}
knitr::include_graphics(pdf2png(here("graphs/02_non_algal_absorption_spectra_vs_fitted.pdf")))
```

## Preliminary analysis

### Comparing aphy vs. chla

This section explore the relationship between aphy and chla and compare it with existing published models.

```{r}
knitr::include_graphics(pdf2png(here("graphs/fig02.pdf")))
```

### Seasonal cycle

- For these graphs, have calculated the year day average of each variable. Then, I have used a 8-days (points) rollmean (red curves).

- I have only used the two bioregions at this time.

- I am not sure if this is really interesting because there are many gaps in the year days, so using a 8 points moving window is not really 8 days.

```{r}
knitr::include_graphics(pdf2png(here("graphs/fig03.pdf")))
```

## TODOS

- Re-do the analysis with only measurements made at the surface layer (<= 2 meters).

- Combine all the data into a single file.

- Add figure numbers.

- Aphy vs chla : cool de voir que des modeles vieux de 15, 20 ans fonctionne toujours, ca serait bien d’avoir les erreurs relatives et absolues de chaques models, ca sera bien pour la discussion

- Time series , je mettrais plutôt « seasonal cycle » and I would remove august and september, there are not enough data.

- Regarder les variations aph443 vs anap443 et aph443/aph443 vs Snap par bioregions et saison

- Est-ce que ces variations changes avec les pigments ? surtout fucox, perid, phae (je vais en trouver d’autres ensuites)

- Je vais reflechir a un moyen d’exploiter toute les longueurs d’onde du phyto, ca serait domange de s’arreter a suelement une !

- Regarder si on voit des tendances a plus long terme (peut-etre des analyses en utilisant gam) avec aph, anap et le rapport aph/anap

## References
