---
title: "NetCOLOR project"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    highlight: zenburn
    css: "theme.css"
    toc: true
    toc_float:
      collapsed: false
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "center",
  cache = TRUE
)

library(here)
library(tidyverse)
library(kableExtra)

library(ggpmthemes)

# Set default ggplot2 font size and font family
# devtools::install_github("PMassicotte/ggpmthemes")
theme_set(theme_light_modified(
  base_family = "Montserrat Alternates",
  base_size = 10
))

theme_update(
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  strip.background = element_blank(),
  strip.text = element_text(face = "bold", size = 10, color = "#4c4a4a")
)

source(here("R","zzz.R"))
```

```{r open_data}
hplc <- vroom::vroom(here("data","clean","hplc.csv"))
metadata <- vroom::vroom(here("data","clean","metadata.csv"))
absorption <- vroom::vroom(here("data","clean","absorption.csv"))
```

This section shows some graphics about the processed data. The data consists in three `.csv` files:

1.  `metadata.csv`

2.  `hplc.csv`

3.  `absorption.csv`

`sample_id` *438116* was duplicated and removed from the data.

## Notes

-   Will be using `depth` instead of `pressure` to characterize the depth of the measurements.

-   Will be only using observations measured in the first 2 meters of the water column. All the following graphs will be using data measured between 0-2 meters.

```{r}
knitr::include_graphics(pdf2png(here("graphs","05_barplot_measurement_depth.pdf")))
```

## Mismatch between metadata, HPLC and absorption

Now, this table shows the metadata that have no HPLC values associated.

```{r}
metadata %>% 
  anti_join(hplc, by = "sample_id") %>% 
  rmarkdown::paged_table()
```

These are the absorption spectra with no HPLC association.

```{r}
absorption %>% 
  anti_join(hplc, by = "sample_id") %>% 
  distinct(sample_id, .keep_all = TRUE) %>% 
  rmarkdown::paged_table()
```

## First visualizations

Most of the data is from the Scotian Shelf, Labrador Sea and Scotian Shelf and Slope.

```{r}
knitr::include_graphics(pdf2png(here("graphs","06_number_measurements_cruise.pdf")))
```

The number of observation is also relatively stable through the years.

```{r}
knitr::include_graphics(pdf2png(here("graphs","06_number_measurements_per_year.pdf")))
```

## Bioregions

Observations were classified into bioregions based on the following criteria (where `position` is `south` if `latitude < 48`):

```{r, eval=FALSE, echo=TRUE}
bathymetry > -300 & position == "North" ~ "Labrador & Greenland Shelves (LGS)"
bathymetry <= -300 & position == "North" ~ "Labrador Sea Bassin (LSB)"
bathymetry >= -300 & position == "South" & yday <= 180 ~ "Scotian Shelf Spring (SSSp)"
bathymetry >= -300 & position == "South" & yday > 180 ~ "Scotian Shelf Fall (SSFa)"
bathymetry < -300 & position == "South"~ "Northwest Atlantic Bassin ocean (NAB)"
```

```{r, fig.align='center'}
knitr::include_graphics(pdf2png(here("graphs","04_number_observations_per_bioregion.pdf")))
```

## HPLC

-   I also noticed there are a lot of `0` in the HPLC data. After discussions, they are true `0` and not missing values.

```{r}
hplc %>%
  select(-abs_vol_l) %>%
  pivot_longer(chlb:zea) %>%
  drop_na() %>%
  mutate(is_zero = value == 0) %>% 
  count(name, is_zero) %>%
  ggplot(aes(x = n, y = name, fill = is_zero)) +
  geom_col() +
  paletteer::scale_fill_paletteer_d("ggthemes::wsj_colors6") +
  labs(
    y = NULL,
    fill = "Value\nis zero"
  ) 
```

-   Some HPLC `sample_id` were not numeric (ex.: FL002), I removed them from the data.

-   There are `hplchla` and `hplcchla` variables in the data. I have merge both columns into one because I suspect it is an error and should be the same variable.

-   `but19` was summed as `but19` + `butlike`.

-   `hex19` was summed as `hex19` + `hexlike` + `hexlike2`.

-   I have calculated `aphy_specific` using `hplcchla`.

-   Here is an attempt to visualize the seasonal cycle of few pigments. I have calculated the average pigment concentration and then the average total concentration to finally calculate the relative average contribution of each pigment.

```{r}
knitr::include_graphics(pdf2png(here("graphs","10_hplc_month_bioregion.pdf")))
```

## Absorption

-   Here is how I named absorption data:

    -   `anap` is the data contained in files with `detritus` in their name.
    -   `ap` is the data contained in files with `particulate` in their name.
    -   `aphy` is the data contained in files with `phytoplankton` in their name.

-   Absorption spectra with any values \<= 0 between 350 and 400 nm have been removed.

-   Absorption spectra with `aphy(440)` \< `aphy(410)` have been removed (possible problem with pigment extraction).

#### ap

```{r}
knitr::include_graphics(pdf2png(here("graphs","07_ap_spectral_profiles_per_bioregion.pdf")))
```

#### aphy

```{r}
knitr::include_graphics(pdf2png(here("graphs","07_aphy_spectral_profiles_per_bioregion.pdf")))
```

#### aphy specific

```{r}
knitr::include_graphics(pdf2png(here("graphs","07_aphy_specific_spectral_profiles_per_bioregion.pdf")))
```

We can observe that there is a specific absorption gradient among bioregions.

```{r}
knitr::include_graphics(pdf2png(here("graphs","07_sina_plot_aphy_specific_443nm.pdf")))
```

#### anap

```{r}
knitr::include_graphics(pdf2png(here("graphs","07_anap_spectral_profiles_per_bioregion.pdf")))
```

### Calculating snap

> The fit was done for data between 380 and 730 nm, excluding the 400--480 and 620--710 nm ranges to avoid any residual pigment absorption that might still have been present after sodium hypochlorite treatment [@Babin2003].

A R2 of 0.90 was used to filter out bad models.

```{r}
knitr::include_graphics(pdf2png(here("graphs","02_non_algal_absorption_spectra_vs_fitted.pdf")))
```

## Preliminary analysis

### Comparing aphy vs. chla

This section explore the relationship between aphy and chla and compare it with existing published models globally and by bioregion.

```{r}
knitr::include_graphics(pdf2png(here("graphs","fig02.pdf")))
```

```{r}
knitr::include_graphics(pdf2png(here("graphs","fig02a.pdf")))
```

### Seasonal cycle

-   The next figure shows how many observations there are for each year of the day. We can see that there are temporal gaps in the data. For example, in `Labrador & Greenland Shelves (LGS)`, there are only data in May/June. It is not enough to get insights on the seasonal cycles. At the moment, I am not sure how to deal with that.

```{r}
knitr::include_graphics(pdf2png(here("graphs","08_number_observations_per_yday_bioregion.pdf")))
```

-   A first attempt can be by using boxplots to see if there are differences among months and bioregions.

```{r}
knitr::include_graphics(pdf2png(here("graphs","08_boxplot_aphy_specific_by_month_bioregion.pdf")))
```

-   Another way to visualize the same data.

```{r}
knitr::include_graphics(pdf2png(here("graphs","08_boxplot_aphy_specific_by_month_bioregion2.pdf")))
```

-   We can also look at the seasonal variability by bioregion. There are interesting patterns in the data. For instance, in `Northwest Atlantic Bassin ocean (NAB)` we can clearly see the seasonal pattern of `hex19`.

```{r}
knitr::include_graphics(pdf2png(here("graphs","08_boxplot_hex19_by_month_bioregion.pdf")))
```

### Phytoplankton population model

Explore and try to understand the method of Emmanuel that aims to decompose absorption spectra into two populations.

```{r}
knitr::include_graphics(pdf2png(here("graphs","11_hplc_chla_vs_phyto_absorption.pdf")))
```

```{r}
knitr::include_graphics(pdf2png(here("graphs","11_s_vs_wavelength.pdf")))
```

```{r}
knitr::include_graphics(pdf2png(here("graphs","11_as1_as2_vs_wavelength.pdf")))
```

### Bricaud 1995

In this section, I put our data in the context of the paper by @Bricaud1995.

-   In table 2, they provide coefficients to fit $a^*_\phi$ as a function of chlorophyll-a between 400 and 700 nm. The next figure shows some randomly selected spectral profiles of $a^*_\phi$ and compare them with those modeled using the coefficients presented in @Bricaud1995.

```{r}
knitr::include_graphics(pdf2png(here("graphs","12_specific_phytoplankton_absorption_spectra_vs_fitted.pdf")))
```

-   The next figure shows $a^*_\phi$ modeled at different wavelengths using the following equation:

$$
a^*_\phi(\lambda) = A(\lambda)\times\text{chl-a}^{-B(\lambda)}
$$

```{r}
knitr::include_graphics(pdf2png(here("graphs","12_specific_phytoplankton_absorption_vs_chla_bricaud_1995_figure01.pdf")))
```

## Fucoxanthin bining

```{r}
knitr::include_graphics(pdf2png(here(
  "graphs", "13_boxplot_percentage_fucoxanthin.pdf"
)))
```

```{r}
knitr::include_graphics(pdf2png(
  here(
    "graphs",
    "13_barplot_number_observations_bining_fucoxanthin.pdf"
  )
))
```

```{r}
knitr::include_graphics(pdf2png(
  here(
    "graphs",
    "13_mean_specific_phytoplankton_absorption_bining_fucoxanthin.pdf"
  )
))
```

### Relationships between fucoxanthin relative content and absorption

```{r}
knitr::include_graphics(pdf2png(
  here(
    "graphs",
    "13_scatterplot_percent_fucoxanthin_aphy_specific.pdf"
  )
))
```

Based on the *linear* relationships in the previous graph, this shows the R2 as a function of the wavelength.

```{r}
knitr::include_graphics(pdf2png(
  here("graphs", "13_r2_percent_fucoxanthin_vs_aphy_specific.pdf")
))
```

## Apparent visible wavelength (AVW)

In this section we explore how the Apparent visible wavelength (AVW) could be used to describe the spectral shape of phytoplankton absorption. The AVW index is calculated following [@vandermeulen2020].

```{r}
knitr::include_graphics(pdf2png(
  here("graphs", "14_boxplot_avw_by_bioregions.pdf")
))
```

This graph shows all the phytoplankton absorption spectra colored by the value of the calculated AVW.

```{r}
knitr::include_graphics(pdf2png(
  here("graphs", "14_normalized_aphy_avw_colored.pdf")
))
```

```{r}
knitr::include_graphics(pdf2png(
  here("graphs", "14_aphy_spectra_highest_avw_difference.pdf")
))
```

```{r}
knitr::include_graphics(pdf2png(
  here("graphs", "14_scatterplot_chla_vs_avw.pdf")
))
```

## Random forest (RF) classification tree

I have used a random forest classification tree to classify the three bioregions using the variables presented in the graphic below. The model performs well with a **ROC AUC value of 0.846**.

> Generally, an ROC AUC value is between 0.5 and 1, with 1 being a perfect prediction model. If your value is between 0 and 0.5, then this implies that you have meaningful information in your model, but it is being applied incorrectly because doing the opposite of what the model predicts would result in an AUC >0.5.

The variables are ranked as a function of their importance to correctly classify the bioregions. For example, within this set of variables, `fucox` is the most important variable whereas `but19` is the least important.

```{r}
knitr::include_graphics(pdf2png(
  here("graphs","17_variable_importance_plot_random_forest.pdf")
))
```


## Figures for the manuscript

### Figure 1

```{r, fig.align='center'}
knitr::include_graphics(pdf2png(here("graphs","fig01.pdf")))
```

### Figure 2

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig02.pdf")))
```

### Figure 3

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig03.pdf")))
```

### Figure 4

There are some unusual bumps around 530 nm. We should have a closer look at the raw data.

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig04.pdf")))
```

### Figure 5

-   Figure 5 : Variations saisonniere de a*p(443) et a*p(670) pour les 4 biomes, on peut avoir 4 panneaux (un pour chaque biome).

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig05.pdf")))
```

### Figure 6

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig06.pdf")))
```

### Figure 7

There are quite many gaps in the data. It was therefore impossible to go lag-correlation because one need one observation per day. To do it so, I took the LOESS models calculated in Fig. 5. Then, I randomly generated data (one observation per day) within the standard error intervals found in Fig. 5. I repeated the process 1000 times (see Appendix 1). After that, I calculated the auto-correlation with a max gap of +- 30 days for each of the simulation. Finally, I calculated the average correlation per lag over the 1000 simulation. The next graphs show the lag-correlation between various absorption index. 

- Panels **A**: Relationships between raw data at lag 0.
- Panels **B**: Lag plots.

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig07.pdf")))
```

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig07a.pdf")))
```

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig07b.pdf")))
```

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig07c.pdf")))
```

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig07d.pdf")))
```

### Figure 8

Here we explore the relationships between the nutrients and the apparent visible wavelength (AVW).

- First, have a look at pairwise scatterplots for the most correlated nutrients to AVW.

```{r}
knitr::include_graphics(pdf2png(here(
  "graphs", "fig08_nutrient_correlation_to_avw.pdf"
)))
```

- Then, I removed the nutrients that were autocorrelated and made a PCA with the others. 

```{r}
knitr::include_graphics(pdf2png(here(
  "graphs", "fig08_pca_avw_nutrients.pdf"
)))
```

- We can also see here that when the spectral shape move toward the blue, there is an exponential increase in fucoxanthin concentration.

```{r}
knitr::include_graphics(pdf2png(here("graphs", "fig08.pdf")))
```

## Appendix

### Appendix 1

**(A)** Simulated examples of `aphy*(443)` and `anap(443)` at a one-day temporal resolution. The black and red lines show the LOESS model and the 99% confidence interval as they appear in Fig. 5. **(B)** Scatterplots showing the relationships between simulated `aphy*(443)` and `anap(443)` at lag 0. Data from Northwest Atlantic Bassin ocean (NAB).

```{r}
knitr::include_graphics(pdf2png(here("graphs", "appendix1.pdf")))
```

## TODOS

-   Is there a variable that can be used to distinguish the stations other than the lat/long variables? I could use this to look at temporal/seasonal patterns in the data.

-   Add figure numbers.

-   Aphy vs chla : cool de voir que des modeles vieux de 15, 20 ans fonctionne toujours, ca serait bien d'avoir les erreurs relatives et absolues de chaques models, ca sera bien pour la discussion

-   Time series , je mettrais plutôt « seasonal cycle » and I would remove august and september, there are not enough data.

-   Regarder les variations aph443 vs anap443 et aph443/aph443 vs Snap par bioregions et saison

-   Est-ce que ces variations changes avec les pigments ? surtout fucox, perid, phae (je vais en trouver d'autres ensuites)

-   Je vais reflechir a un moyen d'exploiter toute les longueurs d'onde du phyto, ca serait domange de s'arreter a suelement une !

    -   We could try the spectral slope curve from @Helms2008.

-   Regarder si on voit des tendances a plus long terme (peut-etre des analyses en utilisant gam) avec aph, anap et le rapport aph/anap

-   What are the units for the pigments?

-   Predict aphy(443) using the main pigments per bioregion and then use variance partitioning to see if there are differences.

## References
