---
title: "NetCOLOR project"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    highlight: zenburn
    css: "theme.css"
    toc: true
    toc_float:
      collapsed: false
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "center",
  cache = TRUE
)

library(here)
library(tidyverse)
library(kableExtra)

library(ggpmthemes)

# Set default ggplot2 font size and font family
# devtools::install_github("PMassicotte/ggpmthemes")
theme_set(theme_light_modified(
  base_family = "Montserrat Alternates",
  base_size = 10
))

theme_update(
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  strip.background = element_blank(),
  strip.text = element_text(face = "bold", size = 10, color = "#4c4a4a")
)

source(here("R/zzz.R"))
```

```{r open_data}
hplc <- vroom::vroom(here("data/clean/hplc.csv"))
metadata <- vroom::vroom(here("data/clean/metadata.csv"))
absorption <- vroom::vroom(here("data/clean/absorption.csv"))
```

This section shows some graphics about the processed data. The data consists in three `.csv` files:

1.  `metadata.csv`

2.  `hplc.csv`

3.  `absorption.csv`

`sample_id` *438116* was duplicated and removed from the data.

## Mismatch between metadata, HPLC and absorption

Now, this table shows the metadata that have no HPLC values associated.

```{r}
metadata %>% 
  anti_join(hplc, by = "sample_id") %>% 
  rmarkdown::paged_table()
```

These are the absorption spectra with no HPLC association.

```{r}
absorption %>% 
  anti_join(hplc, by = "sample_id") %>% 
  distinct(sample_id, .keep_all = TRUE) %>% 
  rmarkdown::paged_table()
```

## First visualizations

Most of the data is from the Scotian Shelf, Labrador Sea and Scotian Shelf and Slope.

```{r}
knitr::include_graphics(pdf2png(here("graphs/06_number_measurements_cruise.pdf")))
```

The number of observation is also relatively stable through the years.

```{r}
knitr::include_graphics(pdf2png(here("graphs/06_number_measurements_per_year.pdf")))
```

## Bioregions

Observations were classified into bioregions based on the following criteria (where `position` is `south` if `latitude < 48`):

```{r, eval=FALSE, echo=TRUE}
bathymetry > -300 & position == "North" ~ "Labrador & Greenland Shelves (LGS)"
bathymetry <= -300 & position == "North" ~ "Labrador Sea Bassin (LSB)"
bathymetry >= -300 & position == "South" & yday <= 180 ~ "Scotian Shelf Spring (SSSp)"
bathymetry >= -300 & position == "South" & yday > 180 ~ "Scotian Shelf Fall (SSFa)"
bathymetry < -300 & position == "South"~ "Northwest Atlantic Bassin ocean (NAB)"
```

```{r, fig.align='center'}
knitr::include_graphics(pdf2png(here("graphs/04_number_observations_per_bioregion.pdf")))
```

## Depth of measurements

- There are two columns in the data that can be used to characterize the depth of the measurement. Which one should be used?

```{r}
df_viz <- absorption %>% 
  distinct(sample_id, depth, pressure)

df_viz %>% 
  drop_na() %>% 
  ggplot(aes(x = depth, y = pressure)) +
  geom_point() +
  geom_abline(lty = 2, color = "red")
```

- Looks like that most measurements were made at `depth == 1`. Should we only use these observations?

```{r}
df_viz %>% 
  count(depth) %>% 
  drop_na() %>% 
  # mutate(depth = fct_reorder(as.character(depth), n))
  ggplot(aes(x = n, y = fct_rev(factor(depth)))) +
  geom_col() +
  labs(
    x = "Number of observations",
    y = "Depth of measurements (m)"
  )
```


## Geographical map

This is not the final map, but it gives a first spatial overview of the data.

```{r, fig.align='center'}
knitr::include_graphics(pdf2png(here("graphs/fig01.pdf")))
```

## HPLC

-   I also noticed there are a lot of `0` in the HPLC data.
    -   Should we consider `0` values as real `0` or missing values?

```{r}
hplc %>%
  select(-abs_vol_l) %>%
  pivot_longer(chlb:zea) %>%
  drop_na() %>%
  mutate(is_zero = value == 0) %>% 
  count(name, is_zero) %>%
  ggplot(aes(x = n, y = name, fill = is_zero)) +
  geom_col() +
  paletteer::scale_fill_paletteer_d("ggthemes::wsj_colors6") +
  labs(
    y = NULL,
    fill = "Value\nis zero"
  ) 
```

- Some HPLC `sample_id` were not numeric (ex.: FL002), I removed them from the data.

-   There are `hplchla` and `hplcchla` variables in the data. I have merge both columns into one because I suspect it is an error and should be the same variable.

- I have calculated `aphy_specific` using `hplcchla`.

## Absorption

-   Here is how I named absorption data:

    -   `anap` is the data contained in files with `detritus` in their name.
    -   `ap` is the data contained in files with `particulate` in their name.
    -   `aphy` is the data contained in files with `phytoplankton` in their name.

-   Is the absorption data ready to go (i.e. no further processing required)?

A quick look at one sample (378972) seems to indicate no...

```{r}
absorption %>% 
  filter(sample_id == 378972) %>%
  ggplot(aes(x = wavelength, y = anap)) +
  geom_line()
```

-   How should we use the quality flags?

-   I think we should only keep the 2671 spectra where all the three flags are at 1.

```{r}
absorption %>%
  filter(wavelength == 350) %>%
  count(
    qc_flag_410_smaller_than_440,
    qc_flag_443_to_490,
    qc_flag_chl_a_to_443_550_670
  ) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_material()
```

Overview pf `aphy` per bioregion. We can see that there are some negative values at low wavelengths (see `Northwest Atlantic Bassin ocean (NAB)`).

```{r}
knitr::include_graphics(pdf2png(here("graphs/07_aph_spectral_profiles_per_bioregion.pdf")))
```

### Calculating snap

> The fit was done for data between 380 and 730 nm, excluding the 400--480 and 620--710 nm ranges to avoid any residual pigment absorption that might still have been present after sodium hypochlorite treatment [@Babin2003].

A R2 of 0.90 was used to filter out bad models.

```{r}
knitr::include_graphics(pdf2png(here("graphs/02_non_algal_absorption_spectra_vs_fitted.pdf")))
```

## Preliminary analysis

### Comparing aphy vs. chla

This section explore the relationship between aphy and chla and compare it with existing published models.

```{r}
knitr::include_graphics(pdf2png(here("graphs/fig02.pdf")))
```

### Timeseries

- For these graphs, have calculated the year day average of each variable. Then, I have used a 8-days (points) rollmean (red curves).

- I have only used the two bioregions at this time.

- I am not sure if this is really interesting because there are many gaps in the year days, so using a 8 points moving window is not really 8 days.

```{r}
knitr::include_graphics(pdf2png(here("graphs/fig03.pdf")))
```

## TODOS

- Re-do the analysis with only measurements made at the surface layer (<= 2 meters).

## References
