---
title: "Seasonal and decadal variability in bio-optical properties in the Northwest Atlantic"
# subtitle: "Mentoring opportunity"
author: 
  - "Philippe Massicotte"
date: '2022-01-21 (updated on: `r Sys.Date()`)'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 5, 
  fig.height = 3.5,
  fig.align = "center",
  dev = "png",
  dpi = 300,
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

library(tidyverse)
library(ggpmthemes)
library(here)

theme_set(theme_minimal(base_family = "Montserrat"))

theme_update(
  panel.border = element_blank(),
  axis.ticks = element_blank()
)

source(here("R","zzz.R"))
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#203240",
  header_font_google = google_font("Patrick Hand"),
  text_font_google = google_font("Montserrat", "500", "500i"),
  code_font_google = google_font("Fira Mono"),
  text_font_size = "24px",
  base_font_size = "20px",
  text_color = "#d6bf61",
  background_color = "#203240",
  header_color = "#b25a56", 
  title_slide_text_color = "#d6bf61",
  text_bold_color = "#b25a56",
  text_bold_font_weight = 600, 
  inverse_background_color = "#232323"
)
```

```{css, echo=FALSE}
a {
  color: deepskyblue;
}

.column-left {
  width: 60%;
  float: left;
  padding-top: 1em;
  margin-right: 0em;
}

.column-right {
  width: 40%;
  float: right;
  padding-top: 0em;
}

.inverse h1{
	color: var(--inverse-header-color);
	font-size: 80px;
	/* font-size: "40em"; */
}

.inverse h2{
	color: var(--inverse-header-color);
	font-size: 70px;
	/* font-size: "40em"; */
}
```

## Background

- Many optical indices used to study the spatial and seasonal response of phytoplankton communities to the **underwater lightscape**.

<br>

- Positive relationships between ** $a_{\phi}$ ** and **chlorophyll-*a*** extensively documented (phytoplankton biomass).

<br>

- ** $a_{\phi}^*$ ** vary due to changes in phytoplankton intracellular composition and pigment packaging effect (used to determine phytoplankton size classes and taxonomic groups).

---

## The motivations (1/n)

- Phytoplankton absorption at 440 nm and  675 nm useful to study the phytoplankton dynamics in several aquatic ecosystems:

  - **Do not fully exploiting the spectral information contained in absorption measurements.**

```{r, out.width="45%"}
df <- read_csv(here("data", "clean", "absorption.csv")) %>% 
  filter(sample_id == 439195)

df_peaks <- df %>% 
  filter(wavelength %in% c(440, 675))

df %>%
  ggplot(aes(x = wavelength, y = aphy)) +
  geom_line() +
  geom_point(data = df_peaks, size = 5, color = "red") +
  labs(
    x = "Wavelength (nm)",
    y = quote(a[phi]~(lambda))
  )
```

---

## The motivations (2/n)

- Phytoplankton absorption is becoming routinely measured by in situ by autonomous profilers and **the emergence of hyperspectral satellite**.

- Long-time series at high spatial resolution are difficult to acquire due to financial and logistical constraints (**often limited around the algal spring bloom**).

- There is a need to develop new indices **to syntheses of all the information contained in a phytoplankton absorption spectrum**.

---

## The sampling

.pull-left[
More than **20 years of data** to provide a detailed picture of the seasonal and the decadal variability bio-optical indices in Northwest Atlantic.

More than **3200 observations** of pigments, chlorophyll-*a* and absorption measurements.
]

.pull-right[
```{r, out.width="90%"}
knitr::include_graphics(pdf2png(here("graphs","appendix01.pdf")))
```
]

---

## The bioregions of the Northwest Atlantic

```{r, out.width="55%"}
knitr::include_graphics(pdf2png(here("graphs","fig01.pdf")))
```

---

## Chlorophyll-*a* vs $a_{\phi}$

Agreement with previously reported models in coastal and marine environments.

```{r, out.width="40%"}
knitr::include_graphics(pdf2png(here("graphs","fig04.pdf")))
```

---

## Packaging effect

Negative relations between this ratio and chlorophyll-*a* is an **indication of packaging effect and changes in the intercellular composition of pigments**.

```{r, out.width="90%"}
knitr::include_graphics(pdf2png(here("graphs","fig05.pdf")))
```

---

## Packaging effect (seasonal variability)

- Indications that the cell size of the phytoplankton is significantly higher in the summer compared to the winter.

- Literature: **acclimating to increasing solar irradiance by reducing intracellular chlorophyll-*a* content and increasing carotenoid content**.

```{r, out.width="80%"}
knitr::include_graphics(pdf2png(here("graphs","appendix04.pdf")))
```

---

class: inverse, center, middle

## Please welcome the phytoplankton apparent absorption wavelength (PAAW)

---

## Phytoplankton apparent absorption wavelength (PAAW)

Weighted harmonic mean to **enable equal weighting of the wavelengths in the calculation**.

\begin{equation}
\text{PAAW} = \frac{\sum\limits_{i=\lambda_1}^{\lambda_n} a_\phi(\lambda_i)}{\sum\limits_{i=\lambda_1}^{\lambda_n} \frac{a_\phi(\lambda_i)}{\lambda_i}}
\end{equation}

In units of nanometres, simply **represents the wavelength balance point around which absorption is evenly distributed**.

---

## Phytoplankton apparent absorption wavelength (PAAW)

- Show what it looks like here.

- Maybe a graph showing the relationship between PAAW and fucox vs aphy440 and fucox.

---

## A better index

```{r, out.width="45%"}
knitr::include_graphics(pdf2png(here("graphs","fig08.pdf")))
```

---

## Decadal variability (last 20 years)

**Spring**: shift toward the red part of the spectrum. **Autumn**: shift toward the blue part of the spectrum.

```{r, out.width="60%"}
knitr::include_graphics(pdf2png(here("graphs","fig10.pdf")))
```
